Mixture Modeling Paper Results

### 2017-03-22 PaperEx170308.Rmd
EOD 170321: Functions Fully Consolidated -Need to re-do the table of contents -Need to re-do example 1 -Need to encorporate error handling from Q edits

```{r loadfunctions, echo=F}

@@

#Runs loops of mixture model over different numbers of components and the normal and skew-normal distribution. This function accepts a data frame of the data and identifiers (variables must be entitled data and id respectively). This is the longest function to run and may fail, but can be re-run with adjustments using various options based on the error message output when the function fails. The default values are to do both skew-normal and normal approximation, with up to 5 components and 10 loops per combination of distribution and number of components. It is not reccommended to change these unless the default function fails.
#accepts a dataframe with columns id and data
fitloops<-function(datawithids,loops=10,maxcp=5){
  dists="Both"
  datawithids<-unique(datawithids)
  data<-datawithids$data
  datapts<-length(data)
  norms<-vector("list")
  errors<-matrix(0, nrow=maxcp, ncol=2)
  loopErr = function(data,i,j,family) {
    tryCatch(smsn.mix(data, nu =3, g = i, get.init = TRUE, criteria = TRUE, group = TRUE, family = family, calc.im=FALSE, obs.prob=TRUE, iter.max=5000, kmeans.param=list(iter.max = 50, n.start = 5)),
             error = function(e) {message(paste("Error in loop", j, "of", family, "with", i, "components", sep=" ")); 
               NA}) 
  }

@@

    fitres<-vector("list")
    for (i in 1:maxcp){
      fitres$no[[i]]<-vector("list", 15)
      fitres$sn[[i]]<-vector("list", 15)
      fitres$no[[i]]<-norms$mc[[i]][[nobj$lmin$mc[i]]]
      fitres$sn[[i]]<-norms$ms[[i]][[nobj$lmin$ms[i]]]
    }
    fitres$loops<-loops
    fitres$maxcp<-maxcp
    fitres$datawithids<-datawithids
    fitres$errors<-errors
    return(fitres)
  } else {
    if (which.max(errors)<=maxcp){
      dist<-"Normal"
    } else {
      dist<-"Skew.normal"
    }
    if (ic<=2){
      stop(paste("More than two loops of the mixture modeling algorithm failed to converge for the", dist, "distribution with", ic, "component(s). First ensure that the necessary packages have been loaded correctly. If the packages are loaded correctly, check for extreme outliers (which may be dropped with extreme caution) or violations of skew-normal or normal distribution assumptions."))
    } else {
      stop(paste("More than two loops of the mixture modeling algorithm failed to converge for the", dist, "distribution with", ic, "component(s). Consider choosing a lower number of components (the default is 5) via the maxcp option of the fitloops function."))
    }
  }  
}

#-----------------------------------------------
#[IC2]  bestfits
#-----------------------------------------------
#Determines the optimal number of distributions and type of distributions (as determined by BIC), was well as the other indicated functions. Also creates table of BICs which can be fed into the bicgraph function to visualize which combination of distributions and component numbers is optimal by BIC.

bestfits<-function(fitres){

  maxcp<-fitres$maxcp
  loops<-fitres$loops
  nobj<-vector("list")

@@

  model<-vector("list")
  model$sn2<-fitres$sn[[2]]
  model$snbest<-fitres$sn[[nobj$minind$sn[1]]]
  model$norm2<-fitres$no[[2]]
  model$normbest<-fitres$no[[nobj$minind$no[1]]]
  model$best<-nobj$best
  model$bictab<-data.frame(normal=c(NA)[rep(c(1), times=maxcp)], skew.normal=c(NA)[rep(c(1), times=maxcp)])
  model$bictab$normal<-nobj$no
  model$bictab$skew.normal<-nobj$sn
  componentnames<-vector()
  for (i in 1:maxcp){
    componentnames[i]<-paste(i, " Component(s):  ", sep="")  
  }    
  rownames(model$bictab)<-componentnames
  colnames(model$bictab)<-c("Normal", "Skew-normal")
  model$summary<-data.frame(desc=c(paste(nobj$distp, "with", nobj$minind[nobj$dist][1,1], "components ", sep=" "), paste("Skew-normal with", nobj$minind$sn, "components ", sep=" "), paste("Normal with", nobj$minind$no, "components ", sep=" "), "Skew-normal with 2 components ", "Normal with 2 components "), bic=c(nobj$best$bic, nobj$sn[nobj$minind$sn], nobj$no[nobj$minind$no], nobj$sn[2], nobj$no[2]))
  rownames(model$summary)<-c("Best Overall", "Best Skew-normal  ", "Best Normal", "Two Skew-normal", "Two Normal")
  colnames(model$summary)<-c("Description         ", "BIC  ")

  model$bestdesc<-list(ncomp=nobj$minind[nobj$dist][1,1], dist=nobj$distf)
  model$datawithids<-fitres$datawithids

  print(paste ("The best model by BIC is", nobj$distp, "with", nobj$minind[nobj$dist][1,1], "components", sep = " "))
  print("")
  print("Bayesian Information Criterion (BIC) Matrix")
  print("BIC should be minimized and a difference of 10 BIC indicates strong evidence that the model with lower BIC is superior")
  print(model$bictab)
  print("Below is a table of the BIC's of the most common distribution and number of component combinations to base a cutpoint on")
  print(model$summary)

@@

@@

#-----------------------------------------------
#[IC3]  bicgraph
#-----------------------------------------------
#Plots BIC to show which combination of distributions and number of components is optimal by BIC. BIC is a criterion which should be minimized (ie we are looking for the most negative value), and a decrease in BIC of 10 or more is "strong evidence" that the model with lower BIC is superior. However in some cases you may want to choose a different distribution and this graph allows for an easy compairison of relative BIC.

bicgraph<-function(model,title="BIC by type and number of components",setcolor=c("magenta", "blue", "black")){
  color<-setcolor
  buffer<-(max(model$bictab)-min(model$bictab))/20
  ylim<-c(min(model$bictab)-buffer, max(model$bictab)+buffer)
  if (is.na(model$bictab$Normal[1])==F) {
    plot(model$bictab$Normal, pch=1, col=color[1], main=title, xlab="Number of components", ylab="BIC", type="o",  xaxt="n", ylim=ylim)
    lines(model$bictab$`Skew-normal`, type="o", pch=2, lty=2, col=color[2])
  } else {
    plot(model$bictab$Normal, type="o", pch=2, lty=2, col=color[2], main=title, xlab="Number of components", ylab="BIC",  xaxt="n", ylim=ylim)
    lines(model$bictab$`Skew-normal`, pch=1, col=color[1], type="o")    
  }  
  points(model$bestdesc$ncomp,model$best$bic,type="o", pch="O", col=color[3],cex=2)
  axis(1, at = seq(1, length(model$bictab$Normal), by = 1))
  suppressWarnings(legend("topright", col=c(color[1], color[2], color[3]), lty=c(1,2,0), pch = c(1, 2, 1), legend=c("Normal", "Skew-normal", "Best by BIC"), cex=c(1,1,1), xjust=1, seg.len=3, title="Distribution"))

@@

#-----------------------------------------------
#[IC4]  fitpick
#-----------------------------------------------
#Selects a specific distribution and format so that it can be read into the uncertainty functions. Default picks the optimized BIC function from bestfits, but this can be adjusted if the user chooses otherwise.

fitpick<-function(fitres,bestfits,dist="",ncomp=NA){
  if (dist=="" & is.na(ncomp)==T) {
    dist=bestfits$bestdesc$dist
    ncomp=bestfits$bestdesc$ncomp
  }
  if (ncomp==1) {
    stop("Single component specified so there is no cut-point")
  }

  data=fitres$datawithids$data
  n=ncomp
  if (dist!="Normal" & dist!="Skew-normal") {
    stop("Distribution must be specified as either 'Normal' or 'Skew-normal'")
  } else if ((dist=="Normal" & is.na(fitres$no[[1]]$bic)==T)| (dist=="Skew-normal" & is.na(fitres$sn[[1]]$bic)==T) | ncomp>length(fitres$no) ){
    stop("Must choose a distribution and number of components that were fitted in the fitloops function")
  } else {
    if (dist=="Normal"){
      dist="no"
    } else {
      dist="sn"
    }

@@

    singlefit<-vector("list")
    singlefit<-fitres[[dist]][[ncomp]]

@@

    for (i in 1:n){
      singlefit$mu[i]<-fitres[[dist]][[ncomp]]$mu[cross$o[i]]
      singlefit$sigma2[i]<-fitres[[dist]][[ncomp]]$sigma2[cross$o[i]]
      singlefit$shape[i]<-fitres[[dist]][[ncomp]]$shape[cross$o[i]]
      singlefit$pii[i]<-fitres[[dist]][[ncomp]]$pii[cross$o[i]]
      singlefit$mu[i]<-fitres[[dist]][[ncomp]]$mu[cross$o[i]]
    }

@@

    par<-vector("list")
    par<-data.frame(mean=c(NA)[rep(c(1), times=n)], sd=c(NA)[rep(c(1), times=n)], shape=c(NA)[rep(c(1), times=n)], pii=c(NA)[rep(c(1), times=n)])
    for (i in 1:n){
      par$mean[i]=singlefit$mu[i]
      par$sd[i]=sqrt(singlefit$sigma2[i])
      par$sigma2[i]=singlefit$sigma2[i]
      par$shape[i]=singlefit$shape[i]
      par$pii[i]=singlefit$pii[i]
    }

    parameters<-vector("list")
    parameters$singlefit<-singlefit
    parameters$par<-par
    parameters$datawithids<-fitres$datawithids
    parameters$desc<-vector("list")
    parameters$desc$ncomp<-ncomp
    parameters$desc$dist<-distp
  }

  return(parameters)
}

#-----------------------------------------------
#[IC5]  uncert
#-----------------------------------------------
#Performs uncertainty calculations.

uncert<-function(fitpickobj){

  data<-fitpickobj$datawithids$data
  par<-fitpickobj$par
  fit<-fitpickobj$singlefit
  n=fitpickobj$desc$ncomp
  #Create Uncertainties
  v<-c(seq(min(data),max(data),0.0001))

  nl<-length(v)
  PD<-vector(mode="numeric", length=n)
  PD[1]<-1
  g1<-0
  g2<-0
  Pv.D<-matrix(0, nl, n)
  Pv.DPD<-matrix(0, nl, n)
  PDT<-matrix(0, nl, 1)
  PD.v<-matrix(0, nl, n)
  for (i in 2:n) {
    PD[i]<-par$pii[i]
    PD[1]<-PD[1]-PD[i]
  }
  for (i in 1:n){
    if (fitpickobj$desc$dist=="Normal"){
      Pv.D[,i]<-dnorm(v,mean=par$mean[i], sd=par$sd[i])
    } else {
      Pv.D[,i]<-dsn(v,xi=par$mean[i],omega=par$sd[i],alpha=par$shape[i])
    }
  }
  for (i in 1:n){
    for (j in 1:nl){
      Pv.DPD[j,i]<-Pv.D[j,i]*PD[i]
    }
  }
  PDT<-rowSums (Pv.DPD)
  for (i in 1:n){
    for (j in 1:nl){
      PD.v[j,i] <- Pv.DPD[j,i] / (PDT[j])
    }
  }

  mcluncert<-1-apply(PD.v, 1, max) 
  uncertmat<-PD.v
  restrmat<-cbind(PD.v, v)
  uncertout<-vector("list")
  uncertout$uncertmat<-uncertmat
  uncertout$mcluncert<-mcluncert
  uncertout$restrmat<-restrmat
  uncertout$singlefit<-fit
  uncertout$v<-v
  uncertout$datawithids<-fitpickobj$datawithids
  uncertout$desc<-fitpickobj$desc
  return(uncertout) 
}

#-----------------------------------------------
#[IC6]  multcut
#-----------------------------------------------
#Calculates raw percentages and counts positive and negative for all possible cut points between distributions
multcut<-function(uncertobj){

  PD.v<-uncertobj$uncertmat
  data<-uncertobj$datawithids$data
  n=uncertobj$desc$ncomp
  ncomp=n
  closest<-function(uncertdf,min,max,level){
    uncertdf2 <- uncertdf[ which(uncertdf$v>=min & uncertdf$v<max), ]
    diff.cut<-abs(uncertdf2$uncertainty.v - level)
    min.diffc<-min(diff.cut)
    y<-which(min.diffc==diff.cut)
    #take first one
    z<-min(y)
    cut<-uncertdf2$v[z]  
    return(cut) 
  }
  g1<-vector("list")
  g2<-vector("list")
  uncertdf<-vector("list")
  uncertainty.v<-vector("list")
  cutpoint<-c(0)[rep(c(1), times=ncomp-1)]
  for (j in 1:(ncomp-1)){
    g1[[j]]<-0
    g2[[j]]<-0
    for (i in 1:j){
      g1[[j]]<-g1[[j]]+PD.v[,i]
    }
    for (i in (j+1):n){
      g2[[j]]<-g2[[j]]+PD.v[,i]
    }
    uncertainty.v[[j]]<-1-pmax(g1[[j]], g2[[j]])
  }

  for (j in 1:(ncomp-1)){
    uncertdf[[j]]<-data.frame(v=uncertobj$v, uncertainty.v=uncertainty.v[[j]])
    cutpoint[j]<-closest(uncertdf[[j]],min(data),max(data),1.0)
  }
  class2<-vector("list")
  for (j in 1:(ncomp-1)){
    class2[[j]]<-ifelse(data>=cutpoint[j], "positive", "negative")
  }
  class<- data.frame(matrix(unlist(class2), nrow=uncertobj$singlefit$n, byrow=F),stringsAsFactors=TRUE)
  class$data<-data

@@

  cutobj<-vector("list")
  cutobj$cutpoint<-cutpoint
  cutobj$class<-class
  cutobj$uncertdf<-uncertdf
  cutobj$type<-"Standard"
  cutobj$g1<-g1
  cutobj$g2<-g2
  cutobj$classtab<-classtab
  cutobj$outtab<-outtab
  cutobj$cutnames<-grpname
  #    cutobj$uncertmat<-uncertobj$uncertmat
  #    cutobj$mcluncert<-uncertobj$mcluncert
  #    cutobj$restrmat<-uncertobj$restrmat
  #    cutobj$singlefit<-uncertobj$fit
  #    cutobj$data<-uncertobj$data
  #    cutobj$desc<-uncertobj$desc
  print(outtab)

@@

#-----------------------------------------------
#[IC7]  rawuncertgraph
#-----------------------------------------------
#Displays the overall uncertainty function (which is 1-max(probability of membership to component i)) with the possible cut-points overlaid. Note: Sometimes the cutpoint does not line up exactly with a peak in the uncertainty function, which looks like a flaw, but is not. In a two component setting, the cutpoint will always occur at the peak of the uncertainty but this is not always the case when there are more than two components.
rawuncertgraph<-function(uncertobj,multcutobj,xlab="X Value",xlim=c(NA,NA),suppresslegend=F,setcolor=c(
  "green4",          #1
  "turquoise3",      #2
  "royalblue2",      #3
  "darkorchid2",     #4
  "firebrick3",      #5
  "darkorange2",     #6
  "white",           #7
  "grey",            #8
  "black"            #9
)){

  grobj<-list()
  if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
    grobj$minval<-(1/10)*trunc(10*min(uncertobj$datawithids$data)) 
    grobj$maxval<-max(uncertobj$datawithids$data)
  } else {
    grobj$minval<-xlim[1]
    grobj$maxval<-xlim[2]
  }

  plot(uncertobj$v, uncertobj$mcluncert, main=paste("Uncertainty Plot with Possible Cutpoints for", uncertobj$desc$dist, "with", uncertobj$desc$ncomp, "components", sep = " "), type="l", ylim=c(0,1), xlim=c(grobj$minval,grobj$maxval), xlab=xlab, ylab="Uncertainty")

  col<-setcolor
  ncomp<-uncertobj$desc$ncomp
  for (i in 1:(uncertobj$desc$ncomp-1)){
    abline(v =multcutobj$cutpoint[i], untf = T, col=col[i], lwd=2)
  }

  if (suppresslegend==F) {
    legtxt<-vector("character", ncomp)
    colors<-vector("character", ncomp)
    for (i in 2:ncomp){
      legtxt[i]<-paste(multcutobj$cutnames[i-1], " , x=", round(multcutobj$cutpoint[i-1], 2), sep="")
      colors[i]<-col[i-1]
    }  
    legtxt[1]<-"Uncertainty"
    colors[1]<-col[length(col)]
    linewidth<-c(2)[rep(c(1), times=ncomp)]
    legend("topright", col=colors, lwd=c(1,linewidth), legend=legtxt, xjust=1, seg.len=3)
  }
}

#-----------------------------------------------
#[IC8]  rawdistgraph
#-----------------------------------------------
#Displays the distributions of all the components for the chosen combination of distribution and number of components. 
rawdistgraph<-function(pickobj,multcutobj,xlim=c(NA,NA),xlab="Optical Density",setbreaks=100,suppresslegend=F,setcolor=c(
  "green4",          #1
  "turquoise3",      #2
  "royalblue2",      #3
  "darkorchid2",     #4
  "firebrick3",      #5
  "darkorange2",     #6
  "white",           #7
  "grey",            #8
  "black"            #9
)){
  #Parameter specification
  ncomp=pickobj$desc$ncomp
  grobj<-vector("list")
  grobj$gpar<-vector("list")
  data<-pickobj$datawithids$data
  color<-setcolor

  if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
    grobj$minval<-(1/10)*trunc(10*min(pickobj$datawithids$data)) 
    grobj$maxval<-max(data)
  } else {
    grobj$minval<-xlim[1]
    grobj$maxval<-xlim[2]
  }

@@

#-----------------------------------------------
#[IC9]  rawhistcuts
#-----------------------------------------------
#Displays which component each bar of the histogram would be assigned to.

>>>>

```{r loadfunctions, echo=F}
#Master Cutoff Algorithm Functions

#The cutoff algorithm contained in this R script allows the user to descriminate between a positive and negative sample on a continuous diagnostic without a good gold standard via mixture modeling.
#The algorithm runs 10 different potential types of mixture models which are based on the normal and skew-normal distributions, looking at 1-5 components each. The optimal distribution and component combination is chosen by BIC (which should be minimized)
#Note: the fitloops function takes some time (typically 2-3 minutes) to run and can fail, if it does, read the error messages and decide whether to try and run it again, or restrict the maximum number of components tested.
#The user then checks which distribution has been chosen and can manually chose a different distribution if desired (there should be good scientific or clinical reasoning to choose a different combination)
#If a model with more than two components is optimal by BIC, then the user must allot components to the positive and negative side using their knowlege of the epidemiological conditions where the data comes from.
#After the decision of where to cut is made, the user can then output graphs and tables that detail what proportion of the population falls into what category, with various degrees of certainty.

#To complete this process, there are 5 necessary steps
#I)   Install/load packages and load in functions (run lines 67 - 1350)
#II)  Load in data (run lines: 1370)
#III) Run each line of code separately, from 1380 to the end of the program, following the directions in the comments
#     The code will: 1)determine which type of distribution and number of sub-populations is optimal; 2) let the user  
#     determine which cutpoint is best (if a model with more than two components is optimal) and 3) display the cutoff results

#-----------------------------------------------
#Table of Contents
#-----------------------------------------------
#[I] Install/load packages and load in functions
#[IA] Install Packages
#[IB] Load Packages
#[IC] Load Functions
#[IC1]  fitloops
#[IC2]  bicgraph
#[IC3]  modelpick
#[IC4]  rawuncertgraph
#[IC5]  rawdistgraph
#[IC6]  rawhistcuts
#[IC7]  cutoff

#-----------------------------------------------
#[I] Install/load packages and load in functions
#-----------------------------------------------

#-----------------------------------------------
#[IA] Install Packages
#-----------------------------------------------

#install.packages("mixsmsn")
#install.packages("sn")

@@

#Runs loops of mixture model over different numbers of components and the normal and skew-normal distribution. This function accepts a data frame of the data and identifiers (variables must be entitled data and id respectively). This is the longest function to run and may fail, but can be re-run with adjustments using various options based on the error message output when the function fails. The default values are to do both skew-normal and normal approximation, with up to 5 components and 10 loops per combination of distribution and number of components. It is not reccommended to change these unless the default function fails.
#accepts a dataframe with columns id and data
fitloops<-function(datawithids,loops=10,maxcp=5){
  dists="Both"
  datawithids<-unique(datawithids)
  data<-datawithids$data
  datapts<-length(data)
  norms<-vector("list")
  errors<-matrix(0, nrow=maxcp, ncol=2)
  loopErr = function(data,i,j,family) {
    tryCatch(suppressWarnings(smsn.mix(data, nu =3, g = i, get.init = TRUE, criteria = TRUE, group = TRUE, family = family, calc.im=FALSE, obs.prob=TRUE, iter.max=5000, kmeans.param=list(iter.max = 50, n.start = 5))),
             error = function(e) {message(paste("Error in loop", j, "of", family, "with", i, "components", sep=" ")); 
               NA}) 
  }

@@

    fitres<-vector("list")
    for (i in 1:maxcp){
      fitres$no[[i]]<-vector("list", 15)
      fitres$sn[[i]]<-vector("list", 15)
      fitres$no[[i]]<-norms$mc[[i]][[nobj$lmin$mc[i]]]
      fitres$sn[[i]]<-norms$ms[[i]][[nobj$lmin$ms[i]]]
    }

    nobj<-vector("list")

@@

    model<-vector("list")

    model$loops<-loops
    model$maxcp<-maxcp
    model$datawithids<-datawithids
    model$errors<-errors
    model$loopfits<-fitres

    model$sn2<-fitres$sn[[2]]
    model$snbest<-fitres$sn[[nobj$minind$sn[1]]]
    model$norm2<-fitres$no[[2]]
    model$normbest<-fitres$no[[nobj$minind$no[1]]]
    model$best<-nobj$best
    model$bictab<-data.frame(normal=c(NA)[rep(c(1), times=maxcp)], skew.normal=c(NA)[rep(c(1), times=maxcp)])
    model$bictab$normal<-nobj$no
    model$bictab$skew.normal<-nobj$sn
    componentnames<-vector()
    for (i in 1:maxcp){
      componentnames[i]<-paste(i, " Component(s):  ", sep="")  
    }    
    rownames(model$bictab)<-componentnames
    colnames(model$bictab)<-c("Normal", "Skew-normal")
    model$summary<-data.frame(desc=c(paste(nobj$distp, "with", nobj$minind[nobj$dist][1,1], "components ", sep=" "), paste("Skew-normal with", nobj$minind$sn, "components ", sep=" "), paste("Normal with", nobj$minind$no, "components ", sep=" "), "Skew-normal with 2 components ", "Normal with 2 components "), bic=c(nobj$best$bic, nobj$sn[nobj$minind$sn], nobj$no[nobj$minind$no], nobj$sn[2], nobj$no[2]))
    rownames(model$summary)<-c("Best Overall", "Best Skew-normal  ", "Best Normal", "Two Skew-normal", "Two Normal")
    colnames(model$summary)<-c("Description         ", "BIC  ")

    model$bestdesc<-list(ncomp=nobj$minind[nobj$dist][1,1], dist=nobj$distf)

    print("")
    print(paste ("The best model by BIC is", nobj$distp, "with", nobj$minind[nobj$dist][1,1], "components", sep = " "))
    print("")
    print("Bayesian Information Criterion (BIC) Matrix")
    print("BIC should be minimized and a difference of 10 BIC indicates strong evidence that the model with lower BIC is superior")
    print(model$bictab)
    print("Below is a table of the BIC's of the most common distribution and number of component combinations to base a cutpoint on")
    print(model$summary)

@@

  } else {
    if (which.max(errors)<=maxcp){
      dist<-"Normal"
    } else {
      dist<-"Skew.normal"
    }
    if (ic<=2){
      stop(paste("More than two loops of the mixture modeling algorithm failed to converge for the", dist, "distribution with", ic, "component(s). First ensure that the necessary packages have been loaded correctly. If the packages are loaded correctly, check for extreme outliers (which may be dropped with extreme caution) or violations of skew-normal or normal distribution assumptions."))
    } else {
      stop(paste("More than two loops of the mixture modeling algorithm failed to converge for the", dist, "distribution with", ic, "component(s). Consider choosing a lower number of components (the default is 5) via the maxcp option of the fitloops function."))
    }
  }

@@

#-----------------------------------------------
#[IC2]  bicgraph
#-----------------------------------------------
#Plots BIC to show which combination of distributions and number of components is optimal by BIC. BIC is a criterion which should be minimized (ie we are looking for the most negative value), and a decrease in BIC of 10 or more is "strong evidence" that the model with lower BIC is superior. However in some cases you may want to choose a different distribution and this graph allows for an easy compairison of relative BIC.

bicgraph<-function(fitobj,title="BIC by type and number of components",setcolor=c("magenta", "blue", "black")){
  model<-fitobj
  color<-setcolor
  buffer<-(max(model$bictab)-min(model$bictab))/20
  ylim<-c(min(model$bictab)-buffer, max(model$bictab)+buffer)
  if (is.na(model$bictab$Normal[1])==F) {
    plot(model$bictab$Normal, pch=1, col=color[1], main=title, xlab="Number of components", ylab="BIC", type="o",  xaxt="n", ylim=ylim)
    lines(model$bictab$`Skew-normal`, type="o", pch=2, lty=2, col=color[2])
  } else {
    plot(model$bictab$Normal, type="o", pch=2, lty=2, col=color[2], main=title, xlab="Number of components", ylab="BIC",  xaxt="n", ylim=ylim)
    lines(model$bictab$`Skew-normal`, pch=1, col=color[1], type="o")    
  }  
  points(model$bestdesc$ncomp,model$best$bic,type="o", pch="O", col=color[3],cex=2)
  axis(1, at = seq(1, length(model$bictab$Normal), by = 1))
  suppressWarnings(legend("topright", col=c(color[1], color[2], color[3]), lty=c(1,2,0), pch = c(1, 2, 1), legend=c("Normal", "Skew-normal", "Best by BIC"), cex=c(1,1,1), xjust=1, seg.len=3, title="Distribution"))

@@

#-----------------------------------------------
#[IC3]  modelpick
#-----------------------------------------------
#Selects a specific distribution and format so that it can be read into the uncertainty functions. Default picks the optimized BIC function from bestfits, but this can be adjusted if the user chooses otherwise. The function also performs uncertainty calculations and calculates raw percentages and counts positive and negative for all possible cut points between distributions

modelpick<-function(fitobj, dist="", ncomp=NA){
  #fitpick<-function(fitres,bestfits,dist="",ncomp=NA){
  fitres<-fitobj
  if (dist=="" & is.na(ncomp)==T) {
    dist=fitres$bestdesc$dist
    ncomp=fitres$bestdesc$ncomp
  }
  if (ncomp==1) {
    stop("Single component specified so there is no cut-point")
  }

  data=fitres$datawithids$data
  n=ncomp
  if (dist!="Normal" & dist!="Skew-normal") {
    stop("Distribution must be specified as either 'Normal' or 'Skew-normal'")
  } else if ((dist=="Normal" & is.na(fitres$loopfits$no[[1]]$bic)==T)| (dist=="Skew-normal" & is.na(fitres$loopfits$sn[[1]]$bic)==T) | ncomp>length(fitres$loopfits$no) ){
    stop("Must choose a distribution and number of components that were fitted in the fitloops function")
  } else {
    if (dist=="Normal"){
      dist="no"
    } else {
      dist="sn"
    }

@@

    singlefit<-vector("list")
    singlefit<-fitres$loopfits[[dist]][[ncomp]]

@@

    for (i in 1:n){
      singlefit$mu[i]<-fitres$loopfits[[dist]][[ncomp]]$mu[cross$o[i]]
      singlefit$sigma2[i]<-fitres$loopfits[[dist]][[ncomp]]$sigma2[cross$o[i]]
      singlefit$shape[i]<-fitres$loopfits[[dist]][[ncomp]]$shape[cross$o[i]]
      singlefit$pii[i]<-fitres$loopfits[[dist]][[ncomp]]$pii[cross$o[i]]
      singlefit$mu[i]<-fitres$loopfits[[dist]][[ncomp]]$mu[cross$o[i]]
    }

@@

    par<-vector("list")
    par<-data.frame(mean=c(NA)[rep(c(1), times=n)], sd=c(NA)[rep(c(1), times=n)], shape=c(NA)[rep(c(1), times=n)], pii=c(NA)[rep(c(1), times=n)])
    for (i in 1:n){
      par$mean[i]=singlefit$mu[i]
      par$sd[i]=sqrt(singlefit$sigma2[i])
      par$sigma2[i]=singlefit$sigma2[i]
      par$shape[i]=singlefit$shape[i]
      par$pii[i]=singlefit$pii[i]
    }
  }

  #Performs uncertainty calculations.
  #uncert<-function(fitpickobj){

  fit<-singlefit
  #Create Uncertainties
  v<-c(seq(min(data),max(data),0.0001))

  nl<-length(v)
  PD<-vector(mode="numeric", length=n)
  PD[1]<-1
  g1<-0
  g2<-0
  Pv.D<-matrix(0, nl, n)
  Pv.DPD<-matrix(0, nl, n)
  PDT<-matrix(0, nl, 1)
  PD.v<-matrix(0, nl, n)
  for (i in 2:n) {
    PD[i]<-par$pii[i]
    PD[1]<-PD[1]-PD[i]
  }
  for (i in 1:n){
    if (dist=="Normal"){
      Pv.D[,i]<-dnorm(v,mean=par$mean[i], sd=par$sd[i])
    } else {
      Pv.D[,i]<-dsn(v,xi=par$mean[i],omega=par$sd[i],alpha=par$shape[i])
    }
  }
  for (i in 1:n){
    for (j in 1:nl){
      Pv.DPD[j,i]<-Pv.D[j,i]*PD[i]
    }
  }
  PDT<-rowSums (Pv.DPD)
  for (i in 1:n){
    for (j in 1:nl){
      PD.v[j,i] <- Pv.DPD[j,i] / (PDT[j])
    }
  }

  mcluncert<-1-apply(PD.v, 1, max) 
  uncertmat<-PD.v
  restrmat<-cbind(PD.v, v)
  uncertout<-vector("list")

  #multcut<-function(uncertobj){

  closest<-function(uncertdf,min,max,level){
    uncertdf2 <- uncertdf[ which(uncertdf$v>=min & uncertdf$v<max), ]
    diff.cut<-abs(uncertdf2$uncertainty.v - level)
    min.diffc<-min(diff.cut)
    y<-which(min.diffc==diff.cut)
    #take first one
    z<-min(y)
    cut<-uncertdf2$v[z]  
    return(cut) 
  }
  g1<-vector("list")
  g2<-vector("list")
  uncertdf<-vector("list")
  uncertainty.v<-vector("list")
  cutpoint<-c(0)[rep(c(1), times=ncomp-1)]
  for (j in 1:(ncomp-1)){
    g1[[j]]<-0
    g2[[j]]<-0
    for (i in 1:j){
      g1[[j]]<-g1[[j]]+PD.v[,i]
    }
    for (i in (j+1):n){
      g2[[j]]<-g2[[j]]+PD.v[,i]
    }
    uncertainty.v[[j]]<-1-pmax(g1[[j]], g2[[j]])
  }

  for (j in 1:(ncomp-1)){
    uncertdf[[j]]<-data.frame(v, uncertainty.v=uncertainty.v[[j]])
    cutpoint[j]<-closest(uncertdf[[j]],min(data),max(data),1.0)
  }
  class2<-vector("list")
  for (j in 1:(ncomp-1)){
    class2[[j]]<-ifelse(data>=cutpoint[j], "positive", "negative")
  }
  class<- data.frame(matrix(unlist(class2), nrow=singlefit$n, byrow=F),stringsAsFactors=TRUE)
  class$data<-data

@@

  cutobj<-vector("list")

  cutobj$singlefit<-singlefit
  cutobj$par<-par
  cutobj$datawithids<-fitres$datawithids
  cutobj$desc<-vector("list")
  cutobj$desc$ncomp<-ncomp
  cutobj$desc$dist<-distp

  cutobj$uncertmat<-uncertmat
  cutobj$mcluncert<-mcluncert
  cutobj$restrmat<-restrmat
  cutobj$v<-v

  cutobj$cutpoint<-cutpoint
  cutobj$class<-class
  cutobj$uncertdf<-uncertdf
  cutobj$type<-"Standard"
  cutobj$g1<-g1
  cutobj$g2<-g2
  cutobj$classtab<-classtab
  cutobj$outtab<-outtab
  cutobj$cutnames<-grpname

  print(outtab)

@@

#-----------------------------------------------
#[IC4]  rawuncertgraph
#-----------------------------------------------
#Displays the overall uncertainty function (which is 1-max(probability of membership to component i)) with the possible cut-points overlaid. Note: Sometimes the cutpoint does not line up exactly with a peak in the uncertainty function, which looks like a flaw, but is not. In a two component setting, the cutpoint will always occur at the peak of the uncertainty but this is not always the case when there are more than two components.
rawuncertgraph<-function(modelpickobj,xlab="X Value",xlim=c(NA,NA),suppresslegend=F,setcolor=c(
  "green4",          #1
  "turquoise3",      #2
  "royalblue2",      #3
  "darkorchid2",     #4
  "firebrick3",      #5
  "darkorange2",     #6
  "white",           #7
  "grey",            #8
  "black"            #9
)){

  grobj<-list()
  if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
    grobj$minval<-(1/10)*trunc(10*min(modelpickobj$datawithids$data)) 
    grobj$maxval<-max(modelpickobj$datawithids$data)
  } else {
    grobj$minval<-xlim[1]
    grobj$maxval<-xlim[2]
  }

  plot(modelpickobj$v, modelpickobj$mcluncert, main=paste("Uncertainty Plot with Possible Cutpoints for", modelpickobj$desc$dist, "with", modelpickobj$desc$ncomp, "components", sep = " "), type="l", ylim=c(0,1), xlim=c(grobj$minval,grobj$maxval), xlab=xlab, ylab="Uncertainty")

  col<-setcolor
  ncomp<-modelpickobj$desc$ncomp
  for (i in 1:(modelpickobj$desc$ncomp-1)){
    abline(v =modelpickobj$cutpoint[i], untf = T, col=col[i], lwd=2)
  }

  if (suppresslegend==F) {
    legtxt<-vector("character", ncomp)
    colors<-vector("character", ncomp)
    for (i in 2:ncomp){
      legtxt[i]<-paste(modelpickobj$cutnames[i-1], " , x=", round(modelpickobj$cutpoint[i-1], 2), sep="")
      colors[i]<-col[i-1]
    }  
    legtxt[1]<-"Uncertainty"
    colors[1]<-col[length(col)]
    linewidth<-c(2)[rep(c(1), times=ncomp)]
    legend("topright", col=colors, lwd=c(1,linewidth), legend=legtxt, xjust=1, seg.len=3)
  }
}

#-----------------------------------------------
#[IC5]  rawdistgraph
#-----------------------------------------------
#Displays the distributions of all the components for the chosen combination of distribution and number of components. 
rawdistgraph<-function(modelpickobj,xlim=c(NA,NA),xlab="Optical Density",setbreaks=100,suppresslegend=F,setcolor=c(
  "green4",          #1
  "turquoise3",      #2
  "royalblue2",      #3
  "darkorchid2",     #4
  "firebrick3",      #5
  "darkorange2",     #6
  "white",           #7
  "grey",            #8
  "black"            #9
)){
  #Parameter specification
  ncomp=modelpickobj$desc$ncomp
  grobj<-vector("list")
  grobj$gpar<-vector("list")
  data<-modelpickobj$datawithids$data
  color<-setcolor

  if (is.na(xlim[1])==T | is.na(xlim[2])==T) {   
    grobj$minval<-(1/10)*trunc(10*min(modelpickobj$datawithids$data)) 
    grobj$maxval<-max(data)
  } else {
    grobj$minval<-xlim[1]
    grobj$maxval<-xlim[2]
  }

@@

#-----------------------------------------------
#[IC6]  rawhistcuts
#-----------------------------------------------
#Displays which component each bar of the histogram would be assigned to.
